---
title: "Disease"
output:
  pdf_document: default
  html_document: default
date: "2024-11-08"
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

```{r packages, include = FALSE}
library(tidyverse)
library(dtplyr)

```

## Choose events of interest

Modify the filter on line 26 so it includes only the events you are interested in.  The fewer events the faster the code runs.  The main function can run off any set of variables that has one real date in it.  So, the grouping is entirely flexible.  The only requirement is that there are ZERO NA values in any of the variables that are used in the groupby.
```{r read_in_data}
#read in file, and limit it to the events of interest
tx<-read_rds('Data/IntermediateFiles/events2.rds')#%>%
  #filter(event_type %in% 'health') #make this filter whatever you are interested in, the fewer the events the faster the code runs



```

```{r get_animal_list}
#create a list of animals
animals<-read_rds('Data/IntermediateFiles/master_animals.rds')


```



## Diesease/Therapy

This chunk adds variables to the original data frame which are new grouping variables based on whatever "gap" you have set. For gap exploration you can run the code multiple times seting new gaps and feeding the previous data frame into the new gap function.  Your original data will get wider and have the flexibility to group by whichever gap setting you choose.

```{r lag_code}
source('FUNCTIONS/fxn_lag_master.R')



arrange_vars <- alist(source_farm, source_state, animal_id, ID, Event, date) #this list should contain all variables of interest including the date variable, do NOT use quotes

sort_vars <- c('source_farm', 'source_state', 'animal_id', 'ID', 'Event') #this list must be identical to the arrange variables but EXCLUDE the date variable and add quotes

#Gap1 : fundamental gap defining a new incidence of disease - in this case it is set to 7 ----------------------

set_dz_gap<- 7 #set at 7 for example, but modify this to whatever makes sense for the question

disease7<-test_fxn1(x = tx%>%rowid_to_column(), #x = your data frame with a rowid variable
                arrange_var = arrange_vars,
                mutate_var = sort_vars,
                prefix = "dz7_",
                gap = set_dz_gap)%>% #gap set to identify regimens
  rename(dz7_key = key, #this variable uniquely identifies the rows of data belonging to each disease, n_distinct can be used to count these, or this variable gan be used to group and summarize drugs used or other things of interest like times touched or locations included (for feet and mastitis)
         dz7_date_gap = date_gap, #this specifies the number of days between this row and the previous row in the disease 
         dz7_ct = lag_ct)%>% #this is a logic variable used primarily for fast data exploration/summary when defining gap setings, it can be summed with no group by function which improves speed
  select(-(contains('lag'))) #this is just to clean up the transient variables used to calculate the disease grouping variables.

#Gap2 : fundamental gap defining a new incidence of disease - in this case it is set to 60----------------------

set_dz_gap<- 60 #set at 7 for example, but modify this to whatever makes sense for the question

disease60<-test_fxn1(x = disease7, #if adding aditional gap options feed in the dataframe created with the previous gap setting
                arrange_var = arrange_vars,
                mutate_var = sort_vars,
                prefix = "dz60_",
                gap = set_dz_gap)%>% #gap set to identify regimens
  rename(dz60_key = key,
         dz60_date_gap = date_gap,
         dz60_ct = lag_ct)%>%
  select(-(contains('lag')))




write_rds(disease60, 'Data/IntermediateFiles/disease60.rds')



```


#quick and dirty comparison of gap settings

```{r}
compare_gaps<-disease60%>%
  group_by(Event)%>%
  summarize(ct_dz7 = sum(dz7_ct), 
            ct_dz60 = sum(dz60_ct))%>%
  ungroup()%>%
  mutate(pct_diff_gap = round(((ct_dz7-ct_dz60)/ct_dz7)*100, digits = 4))

```


```{r outcomes}

#health events animal - each row is an animal  -------------------------
disease_survival_animal_level<-outcomes
  arrange(animal_id, dz_events, date)%>%
  group_by(animal_id, dz_events, dz_key)%>%
  mutate(dz_ct_animal = 1:n(), 
         dz_ct_txt = paste0(dz_events, event_ct_animal),
         date_start = min(date), 
         date_end = max(date))%>%
  ungroup()%>%
  select(animal_id, date_start,  event_ct_animal_txt)%>%
  distinct()%>%
  pivot_wider(names_from = event_ct_animal_txt, 
              values_from = date_start)%>%
  distinct()

write_rds(disease_survival_animal_level, 'DATA/IntermediateFiles/disease_survival_animal_level.rds')

#health events lactation - each row is an animal lactation  -------------------------
disease_survival_animal_lact_level<-outcomes
  arrange(animal_id, animal_id_lact,  dz_events, date)%>%
  group_by(animal_id, animal_id_lact, dz_events, dz_key)%>%
  mutate(dz_ct_animal = 1:n(), 
         dz_ct_txt = paste0(dz_events, event_ct_animal),
         date_start = min(date), 
         date_end = max(date))%>%
  ungroup()%>%
  select(animal_id, animal_id_lact, date_start,  event_ct_animal_txt)%>%
  distinct()%>%
  pivot_wider(names_from = event_ct_animal_txt, 
              values_from = date_start)%>%
  distinct()

write_rds(disease_survival_animal_lact_level, 'DATA/IntermediateFiles/disease_survival_animal_lact_level.rds')



```





We need to add the dead detail data before we can interpret outcomes. They are extremely OPTIMISTIC without it. 

#### 30 day outcomes

```{r outcome_summary}
outcomes30<-df_30%>%
  group_by(Pen_simple, g30_outcome, g30_pull_ct)%>%
  summarize(
            ct_pull = n_distinct(g30_key))%>%
  ungroup()%>%
  pivot_wider(names_from = g30_outcome, values_from = ct_pull)%>%
  mutate(total = `Final Tx` + Retreat)%>%
  mutate(pctRetreat = Retreat/total, 
         pctSuccess = `Final Tx`/total)

targets<-tibble(g30_pull_ct = c(1, 2, 3), 
                target_success = c(.8, .5, .3), 
                limit_retreat = c(.2, .5, .7))

ggplot(outcomes30%>%
         filter(g30_pull_ct<4))+
  geom_hline(data = targets,
             aes(yintercept = target_success), color = 'red')+
    geom_text(data = targets,
              aes(y = target_success, x = Inf, label = 'Target'), color = 'red', hjust = 1, vjust = 1)+
  geom_bar(aes(x = Pen_simple, y = pctSuccess, fill = factor(g30_pull_ct)), stat = 'identity')+
  facet_grid(g30_pull_ct~.)+
  labs(title = 'Treatment Success')+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c('#252525', '#636363', '#969696'))+
  theme_bw()+
  labs(fill = 'Pull Count')+
  ylab('Percent of Treatments')+
  xlab('Pen')


# ggplot(outcomes30%>%
#          filter(g30_pull_ct<4))+
#   geom_bar(aes(x = Pen_simple, y = pctRetreat, fill = factor(g30_pull_ct)), stat = 'identity')+
#   geom_hline(aes(yintercept = .2), color = 'black')+
#     geom_text(aes(y = .2, x = Inf, label = 'Limit'), color = 'black', hjust = 1, vjust = 1)+
# 
#   facet_grid(g30_pull_ct~.)+
#   labs(title = 'Treatment Failures')+
#   scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
#   scale_fill_manual(values = c('#a50f15', '#de2d26', '#fb6a4a'))+
#   theme_bw()



```

\newpage




#### 3 day outcomes

These are the outcomes if we evaluate at day 3 after first treatment.

```{r}
outcomes3<-df_30%>%
  group_by(Pen_simple, g3_outcome, g3_pull_ct)%>%
  summarize(
            ct_pull = n_distinct(g3_key))%>%
  ungroup()%>%
  pivot_wider(names_from = g3_outcome, values_from = ct_pull)%>%
  mutate(total = `Final Tx` + Retreat)%>%
  mutate(pctRetreat = Retreat/total, 
         pctSuccess = `Final Tx`/total)

targets<-tibble(g3_pull_ct = c(1, 2, 3), 
                target_success = c(.8, .5, .3))

ggplot(outcomes3%>%
         filter(g3_pull_ct<4))+
  geom_hline(data = targets,
             aes(yintercept = target_success), color = 'red')+
    geom_text(data = targets,
              aes(y = target_success, x = Inf, label = 'Target'), color = 'red', hjust = 1, vjust = 1)+
  geom_bar(aes(x = Pen_simple, y = pctSuccess, fill = factor(g3_pull_ct)), stat = 'identity')+
  facet_grid(g3_pull_ct~.)+
  labs(title = 'Treatment Success')+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c('#252525', '#636363', '#969696'))+
  theme_bw()+
  labs(fill = 'Pull Count')+
  ylab('Percent of Treatments')+
  xlab('Pen')


# ggplot(outcomes3%>%
#          filter(g3_pull_ct<4))+
#   geom_bar(aes(x = Pen_simple, y = pctRetreat, fill = factor(g3_pull_ct)), stat = 'identity')+
#   geom_hline(aes(yintercept = .2), color = 'black')+
#     geom_text(aes(y = .2, x = Inf, label = 'Limit'), color = 'black', hjust = 1, vjust = 1)+
# 
#   facet_grid(g3_pull_ct~.)+
#   labs(title = 'Treatment Failures')+
#   scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
#   scale_fill_manual(values = c('#a50f15', '#de2d26', '#fb6a4a'))+
#   theme_bw()
# 


```

\newpage

#### 7 day outcomes

These are the outcomes if we evaluate at day 7 after first treatment.

```{r}
outcomes7<-df_30%>%
  group_by(Pen_simple, g7_outcome, g7_pull_ct)%>%
  summarize(
            ct_pull = n_distinct(g7_key))%>%
  ungroup()%>%
  pivot_wider(names_from = g7_outcome, values_from = ct_pull)%>%
  mutate(total = `Final Tx` + Retreat)%>%
  mutate(pctRetreat = Retreat/total, 
         pctSuccess = `Final Tx`/total)

targets<-tibble(g7_pull_ct = c(1, 2, 3), 
                target_success = c(.8, .5, .3))

ggplot(outcomes7%>%
         filter(g7_pull_ct<4))+
  geom_hline(data = targets,
             aes(yintercept = target_success), color = 'red')+
    geom_text(data = targets,
              aes(y = target_success, x = Inf, label = 'Target'), color = 'red', hjust = 1, vjust = 1)+
  geom_bar(aes(x = Pen_simple, y = pctSuccess, fill = factor(g7_pull_ct)), stat = 'identity')+
  facet_grid(g7_pull_ct~.)+
  labs(title = 'Treatment Success')+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c('#252525', '#636363', '#969696'))+
  theme_bw()+
  labs(fill = 'Pull Count')+
  ylab('Percent of Treatments')+
  xlab('Pen')



# ggplot(outcomes7%>%
#          filter(g7_pull_ct<4))+
#   geom_bar(aes(x = Pen_simple, y = pctRetreat, fill = factor(g7_pull_ct)), stat = 'identity')+
#   geom_hline(aes(yintercept = .2), color = 'black')+
#     geom_text(aes(y = .2, x = Inf, label = 'Limit'), color = 'black', hjust = 1, vjust = 1)+
# 
#   facet_grid(g7_pull_ct~.)+
#   labs(title = 'Treatment Failures')+
#   scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
#   scale_fill_manual(values = c('#a50f15', '#de2d26', '#fb6a4a'))+
#   theme_bw()



```



\newpage

 

### Graphical Animal Treatment Details

Here is a quick way to look at individual animal treatment patterns.  These are some examples here, but can explore them all at:
\textcolor{blue}{https://lvr-explore.shinyapps.io/AnimalTreatmentDetails/}. 
 
I will provide a link so you can look at any animal after we add the dead list so we know the final outcomes.
My observation is that our protocols are designed in a way that we touch these animals more often than I am used to seing.  I wonder if we are getting the right balance between enough drug without too much handling? 

In most scenarios I am very comfortable with using a Nuflor protocol that is a single injection at 6 ml per 100lbs with a 4 day post treatment interval.  This results in the same or better outcome with less handling. The abomasal leasions make me unsure that we know what we have here, so that is why am I not strongly recommending this change.

Baytril is a very short acting drug, however it is also a concentration dependent drug.  Therefore it has its best outcome at high dose.  Not sure what dose you are using here (3ml per 100?), but consider moving to 5 ml per 100 or greater if not already there.  This is a very short acting drug so 2 days is a very resonable post treatment interval. It gets very expensive quickly to come back with another full dose of baytril unless we know it is helping significantly.



```{r}

set_animal<-c('33223', '1149', '33222', '7185')

ggplot(tx%>%
         filter(tag %in% set_animal)%>%
         mutate(Protocol = factor(Protocol, levels = levels_tx))
       )+
  geom_point(aes(x = Date, y = Protocol, color = Protocol), size = 5, shape = 16)+
  theme_bw()+
  facet_grid(tag~., scales = 'free_y')+
  scale_color_manual(values = colors_drug)+
  labs(title = 'Individual Drug Treatments')
  
 
  

  
  
ggplot(therapy%>%
         filter(tag %in% set_animal)%>%
         mutate(Therapy = case_when(
           (str_count(Protocols_simple)>35)~'Many Treatments', 
           TRUE~Protocols_simple)
         )
       )+
  geom_linerange(aes(xmin = date_tx_start, xmax = date_tx_end, y = Therapy), linewidth = 3)+
    geom_point(aes(x = date_tx_start,  y = Therapy), size = 5)+
  theme_bw()+
  facet_grid(tag~., scales = 'free')+
  labs(title = 'Therapy Durations')+
  xlab('Date')
  

```


