---
title: "Untitled"
output: html_document
date: "2024-11-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

```

```{r}

events2<-read_rds('events2.rds')%>%
  filter(!(is.na(bdat)))%>% 
  mutate(data_pull_date = max(date, na.rm = TRUE))%>%
  mutate(data_pull_date = max(date, na.rm = TRUE))%>%
  mutate(
    breed = case_when(
      is.na(breed)~'Unkown', 
      TRUE~breed), 
    source_farm = case_when(
      is.na(source_farm)~'Unknown', 
      TRUE~source_farm), 
    source_state = case_when(
      is.na(source_state)~'Unknown',
      TRUE~source_state)
          )

```


## Treatment Outcomes (Incomplete - need dead data)

```{r lag_code}

source('FUNCTIONS/fxn_lag_master.R')




arrange_vars <- alist(tag, eid, date) #removed: Reason2, Disease, we need a "stop if list contains 'lag'

sort_vars <- c('tag', 'eid') #removed:"Reason2", "Disease",

#Gap1 ----------------------

set_outcome_gap<- 1

df_1<-test_fxn1(x = treats%>%
                  mutate(date = Date),
                arrange_var = arrange_vars,
                mutate_var = sort_vars,
                prefix = "g1_",
                gap = set_outcome_gap)%>% #gap set to identify regimens
  rename(g1_key = key,
         g1_date_gap = date_gap,
         g1_ct = lag_ct)%>%
  select(-(contains('lag')))%>%
  group_by(tag, eid, g1_key)%>%
  mutate(g1_pull_ct = 1:n(), 
         g1_pull_ct_max = sum(n()))%>%
  ungroup()%>%
  mutate(g1_outcome = case_when(
    (g1_pull_ct < g1_pull_ct_max)~'Retreat', 
    (g1_pull_ct == g1_pull_ct_max)~'Final Tx', 
    TRUE~'Error')
  )


#Gap2 ----------------------

set_outcome_gap<- 2

df_2<-test_fxn1(x = df_1,
                arrange_var = arrange_vars,
                mutate_var = sort_vars,
                prefix = "g2_",
                gap = set_outcome_gap)%>% #gap set to identify regimens
  rename(g2_key = key,
         g2_date_gap = date_gap,
         g2_ct = lag_ct)%>%
  select(-(contains('lag')))%>%
  group_by(tag, eid, g2_key)%>%
  mutate(g2_pull_ct = 1:n(), 
         g2_pull_ct_max = sum(n()))%>%
  ungroup()%>%
  mutate(g2_outcome = case_when(
    (g2_pull_ct < g2_pull_ct_max)~'Retreat', 
    (g2_pull_ct == g2_pull_ct_max)~'Final Tx', 
    TRUE~'Error')
  )


#Gap3 ----------------------

set_outcome_gap<- 3

df_3<-test_fxn1(x = df_2,
                arrange_var = arrange_vars,
                mutate_var = sort_vars,
                prefix = "g3_",
                gap = set_outcome_gap)%>% #gap set to identify regimens
  rename(g3_key = key,
         g3_date_gap = date_gap,
         g3_ct = lag_ct)%>%
  select(-(contains('lag')))%>%
  group_by(tag, eid, g3_key)%>%
  mutate(g3_pull_ct = 1:n(), 
         g3_pull_ct_max = sum(n()))%>%
  ungroup()%>%
  mutate(g3_outcome = case_when(
    (g3_pull_ct < g3_pull_ct_max)~'Retreat', 
    (g3_pull_ct == g3_pull_ct_max)~'Final Tx', 
    TRUE~'Error')
  )

#Gap7 ----------------------

set_outcome_gap<- 7

df_7<-test_fxn1(x = df_3,
                arrange_var = arrange_vars,
                mutate_var = sort_vars,
                prefix = "g7_",
                gap = set_outcome_gap)%>% #gap set to identify regimens
  rename(g7_key = key,
         g7_date_gap = date_gap,
         g7_ct = lag_ct)%>%
  select(-(contains('lag')))%>%
  group_by(tag, eid, g7_key)%>%
  mutate(g7_pull_ct = 1:n(), 
         g7_pull_ct_max = sum(n()))%>%
  ungroup()%>%
  mutate(g7_outcome = case_when(
    (g7_pull_ct < g7_pull_ct_max)~'Retreat', 
    (g7_pull_ct == g7_pull_ct_max)~'Final Tx', 
    TRUE~'Error')
  )

#Gap30 ----------------------

set_outcome_gap<- 30

df_30<-test_fxn1(x = df_7,
                arrange_var = arrange_vars,
                mutate_var = sort_vars,
                prefix = "g30_",
                gap = set_outcome_gap)%>% #gap set to identify regimens
  rename(g30_key = key,
         g30_date_gap = date_gap,
         g30_ct = lag_ct)%>%
  select(-(contains('lag')))%>%
  group_by(tag, eid, g30_key)%>%
  mutate(g30_pull_ct = 1:n(), 
         g30_pull_ct_max = sum(n()))%>%
  ungroup()%>%
  mutate(g30_outcome = case_when(
    (g30_pull_ct < g30_pull_ct_max)~'Retreat', 
    (g30_pull_ct == g30_pull_ct_max)~'Final Tx', 
    TRUE~'Error')
  )





```




We need to add the dead detail data before we can interpret these. They are extremely OPTIMISTIC without it. 
That said, the treatment failure rate is higher than I would like across the yard, not just the pens we are concerned about even before we add the deads in.

#### 30 day outcomes

```{r outcome_summary}
outcomes30<-df_30%>%
  group_by(Pen_simple, g30_outcome, g30_pull_ct)%>%
  summarize(
            ct_pull = n_distinct(g30_key))%>%
  ungroup()%>%
  pivot_wider(names_from = g30_outcome, values_from = ct_pull)%>%
  mutate(total = `Final Tx` + Retreat)%>%
  mutate(pctRetreat = Retreat/total, 
         pctSuccess = `Final Tx`/total)

targets<-tibble(g30_pull_ct = c(1, 2, 3), 
                target_success = c(.8, .5, .3), 
                limit_retreat = c(.2, .5, .7))

ggplot(outcomes30%>%
         filter(g30_pull_ct<4))+
  geom_hline(data = targets,
             aes(yintercept = target_success), color = 'red')+
    geom_text(data = targets,
              aes(y = target_success, x = Inf, label = 'Target'), color = 'red', hjust = 1, vjust = 1)+
  geom_bar(aes(x = Pen_simple, y = pctSuccess, fill = factor(g30_pull_ct)), stat = 'identity')+
  facet_grid(g30_pull_ct~.)+
  labs(title = 'Treatment Success')+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c('#252525', '#636363', '#969696'))+
  theme_bw()+
  labs(fill = 'Pull Count')+
  ylab('Percent of Treatments')+
  xlab('Pen')


# ggplot(outcomes30%>%
#          filter(g30_pull_ct<4))+
#   geom_bar(aes(x = Pen_simple, y = pctRetreat, fill = factor(g30_pull_ct)), stat = 'identity')+
#   geom_hline(aes(yintercept = .2), color = 'black')+
#     geom_text(aes(y = .2, x = Inf, label = 'Limit'), color = 'black', hjust = 1, vjust = 1)+
# 
#   facet_grid(g30_pull_ct~.)+
#   labs(title = 'Treatment Failures')+
#   scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
#   scale_fill_manual(values = c('#a50f15', '#de2d26', '#fb6a4a'))+
#   theme_bw()



```

\newpage




#### 3 day outcomes

These are the outcomes if we evaluate at day 3 after first treatment.

```{r}
outcomes3<-df_30%>%
  group_by(Pen_simple, g3_outcome, g3_pull_ct)%>%
  summarize(
            ct_pull = n_distinct(g3_key))%>%
  ungroup()%>%
  pivot_wider(names_from = g3_outcome, values_from = ct_pull)%>%
  mutate(total = `Final Tx` + Retreat)%>%
  mutate(pctRetreat = Retreat/total, 
         pctSuccess = `Final Tx`/total)

targets<-tibble(g3_pull_ct = c(1, 2, 3), 
                target_success = c(.8, .5, .3))

ggplot(outcomes3%>%
         filter(g3_pull_ct<4))+
  geom_hline(data = targets,
             aes(yintercept = target_success), color = 'red')+
    geom_text(data = targets,
              aes(y = target_success, x = Inf, label = 'Target'), color = 'red', hjust = 1, vjust = 1)+
  geom_bar(aes(x = Pen_simple, y = pctSuccess, fill = factor(g3_pull_ct)), stat = 'identity')+
  facet_grid(g3_pull_ct~.)+
  labs(title = 'Treatment Success')+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c('#252525', '#636363', '#969696'))+
  theme_bw()+
  labs(fill = 'Pull Count')+
  ylab('Percent of Treatments')+
  xlab('Pen')


# ggplot(outcomes3%>%
#          filter(g3_pull_ct<4))+
#   geom_bar(aes(x = Pen_simple, y = pctRetreat, fill = factor(g3_pull_ct)), stat = 'identity')+
#   geom_hline(aes(yintercept = .2), color = 'black')+
#     geom_text(aes(y = .2, x = Inf, label = 'Limit'), color = 'black', hjust = 1, vjust = 1)+
# 
#   facet_grid(g3_pull_ct~.)+
#   labs(title = 'Treatment Failures')+
#   scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
#   scale_fill_manual(values = c('#a50f15', '#de2d26', '#fb6a4a'))+
#   theme_bw()
# 


```

\newpage

#### 7 day outcomes

These are the outcomes if we evaluate at day 7 after first treatment.

```{r}
outcomes7<-df_30%>%
  group_by(Pen_simple, g7_outcome, g7_pull_ct)%>%
  summarize(
            ct_pull = n_distinct(g7_key))%>%
  ungroup()%>%
  pivot_wider(names_from = g7_outcome, values_from = ct_pull)%>%
  mutate(total = `Final Tx` + Retreat)%>%
  mutate(pctRetreat = Retreat/total, 
         pctSuccess = `Final Tx`/total)

targets<-tibble(g7_pull_ct = c(1, 2, 3), 
                target_success = c(.8, .5, .3))

ggplot(outcomes7%>%
         filter(g7_pull_ct<4))+
  geom_hline(data = targets,
             aes(yintercept = target_success), color = 'red')+
    geom_text(data = targets,
              aes(y = target_success, x = Inf, label = 'Target'), color = 'red', hjust = 1, vjust = 1)+
  geom_bar(aes(x = Pen_simple, y = pctSuccess, fill = factor(g7_pull_ct)), stat = 'identity')+
  facet_grid(g7_pull_ct~.)+
  labs(title = 'Treatment Success')+
  scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
  scale_fill_manual(values = c('#252525', '#636363', '#969696'))+
  theme_bw()+
  labs(fill = 'Pull Count')+
  ylab('Percent of Treatments')+
  xlab('Pen')



# ggplot(outcomes7%>%
#          filter(g7_pull_ct<4))+
#   geom_bar(aes(x = Pen_simple, y = pctRetreat, fill = factor(g7_pull_ct)), stat = 'identity')+
#   geom_hline(aes(yintercept = .2), color = 'black')+
#     geom_text(aes(y = .2, x = Inf, label = 'Limit'), color = 'black', hjust = 1, vjust = 1)+
# 
#   facet_grid(g7_pull_ct~.)+
#   labs(title = 'Treatment Failures')+
#   scale_y_continuous(labels = scales::percent_format(accuracy = 1))+
#   scale_fill_manual(values = c('#a50f15', '#de2d26', '#fb6a4a'))+
#   theme_bw()



```


