---
title: "Explore Outcomes"
output: html_document
date: "2024-08-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#setup
```{r}
library(tidyverse)
source('FUNCTIONS/fxn_sourcefarms.R')

```
Some interesting questions:
1) Calves that are given an ASSIST protocol with a remark of TUL.3.1 were given tulathromycin. This is given at the discretion of my transition managers. It’s not quite metaphylaxis but the guys view these calves as high risk. I’d love to understand how these calves perform though the system


2)First-pull recovery rate. Current we are using enro for first pull. I’d curious to know what our recover rate is

#read in event files
```{r}
list_files<-list.files('KDDEvents')

events<-NULL

for (i in seq_along(list_files)){
  df<-read_csv(paste0('KDDEvents/', list_files[i]), col_types = cols(.default = 'c'))%>% #reads in all data as character string
    mutate(Remark = str_replace_all(Remark, "[^[:alnum:]]", "_")) #gets rid of weird characters that mess up encoding or parsing
    
  events<-bind_rows(events, df)
}

```
#initial cleanup
```{r}

#--------------------------------------------------


events2<-events%>%
  filter(!(is.na(Date)))%>%
  select(-starts_with('...'))%>%
  #basid--------------------------------------- 
  mutate(cowid = paste0(ID, BDAT), 
         breed = CBRD)%>%
  mutate(abdat = lubridate::mdy(ABDAT),
         ardat = lubridate::mdy(ARDAT),
         bdat = lubridate::mdy(BDAT), 
         cdat = lubridate::mdy(CDAT), 
         date = lubridate::mdy(Date), 
         ddat = lubridate::mdy(DDAT),
         edat = lubridate::mdy(EDAT), 
         fdat = lubridate::mdy(FDAT), 
         hdat = lubridate::mdy(HDAT),
         podat = lubridate::mdy(PODAT)
         )%>%
  mutate(DIM = parse_number(DIM))%>%
  arrange(cowid, date)%>%
  distinct()%>%
  
  #define event type------------------------------------
  mutate(EVT_type = case_when(
  Event %in% c("ABORT","BRED", "BULLPEN", "FMARTIN", "GNRH", "HEAT", "MV2AI", 
               "OK",      "OPEN", "PREG", "PREV", "PROST", "RECHK" )~'repro', 
  Event %in% c ("ASSIST", "BLOAT", "DIPTHRA", "FEVER", "HLTHSCR", "ILLMISC", "INDIG", "INJURY", 
                "LAME", "MAST", "NAVEL", "OTHER","OTITIS",  "PINKEYE", "PNEU", "RP", 
                "SCOURS",  "SEPTIC", "SWLNLEG")~'health', 
  Event %in% c("ARVDPOT", "DNB", "DRY", "GOAMISH", "GODEER", "GOHOME",  "MOVE")~'mgmt', 
  Event %in% c("ARVKDD", "BEEF", "DIED", "FRESH", "SOLD")~'param', 
  Event %in% c("INWEIGH", "MEASURE", "TP")~'measure',
  Event %in% c("CHECK", "CHKDBN", "CLFSERU", "COMMENT", "INV",  "OBSERVE", "TBTEST",  
               "TRIAL", "USER6", "USER7",  "XID", "ZPRBRED")~'ask', 
    Event %in% c("BANGVAC", "VACC")~'vac', 
  TRUE~'Unknown'
   ))%>%
  

  #add source farm ------------------
  fxn_add_source_farm()


write_rds(events2, 'events2.rds')

  
```






