---
title: "Explore Outcomes"
output: html_document
date: "2024-08-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

```
Some interesting questions:
1) Calves that are given an ASSIST protocol with a remark of TUL.3.1 were given tulathromycin. This is given at the discretion of my transition managers. It’s not quite metaphylaxis but the guys view these calves as high risk. I’d love to understand how these calves perform though the system


2)First-pull recovery rate. Current we are using enro for first pull. I’d curious to know what our recover rate is


```{r}
#--------------------------------------------------

events2<-read_rds('Data/IntermediateFiles/events2.rds')
  
```





```{r}
event_counts<-events2%>%
  group_by(EVT_type, Event, floor_date_month)%>%
  summarize(ct_cowid = n_distinct(cowid))%>%
  ungroup()

ggplot(event_counts%>%filter(EVT_type %in% 'health')%>%
         filter(!(Event %in% c('RP',  'ILLMISC', 'INJURY', 'INDIG', 'MAST', 'SWLNLEG', 'FEVER', 'DIPTHRA'))))+
  geom_point(aes(x = floor_date_month, y = ct_cowid, color = Event))+
    geom_smooth(aes(x = floor_date_month, y = ct_cowid, color = Event, fill = Event, group = Event))#+

  #facet_wrap(Event~., scales = 'free')


```


```{r}

remarks<-events2%>%
  #filter(Event %in% 'PNEU')%>%
  filter(event_ct==1)%>%
  group_by(Event, floor_date_month)%>%
  mutate(total_cowid = n_distinct(cowid), 
            total_row = sum(n()))%>%
  ungroup()%>%
  mutate(Remark_simple = str_sub(Remark, 1,3), collapse = ',')%>%
  mutate(
    draxxin = case_when(
      str_detect(Remark, 'DRA.|DRX.')~'tulathromycin', 
      TRUE~''), 
    baytril = case_when(
      str_detect(Remark, 'BAY.|ENR.')~'enrofloxacin', 
      TRUE~''), 
    excenel = case_when(
      str_detect(Remark, 'EXN.|EXL.|XNL.')~'ceftiofur hcl', 
      TRUE~''), 
    ampicillin = case_when(
      str_detect(Remark, 'AMP.|PF.|PLY.|POL.')~'ampicillin', 
      TRUE~''), 
    excede = case_when(
      str_detect(Remark, 'EXD.|EXC.')~'ceftiofur cfa', 
      TRUE~''), 
    nuflor = case_when(
      str_detect(Remark, 'NU.|RES.|NOR.')~'florfenicol', 
      TRUE~''), 
    zuprevo = case_when(
      str_detect(Remark, 'ZUP.')~'tildipirison', 
      TRUE~''),
    zactran = case_when(
      str_detect(Remark, 'ZAC.')~'gamithromycin', 
      TRUE~''), 
    oxytet = case_when(
      str_detect(Remark, 'OXY.|LA2.|TET.|LAT.|LA.')~'oxytetracycline', 
      TRUE~''), 
    sulf = case_when(
      str_detect(Remark, 'SMZ.')~'sulfa', 
      TRUE~''),
    pen = case_when(
      str_detect(Remark, 'PEN.')~'penicillin', 
      TRUE~''), 
    elect = case_when(
      str_detect(Remark, 'ELEC.|ELE.|LRS.')~'electrolytes', 
      TRUE~''), 
    ban = case_when(
      str_detect(Remark, 'BAN.')~'flunixin', 
      TRUE~''), 
    thera = case_when(
      str_detect(Remark, 'TH.')~'therabloat', 
      TRUE~''),
    dex = case_when(
      str_detect(Remark, 'DEX.|DEX')~'dexamethazone', 
      TRUE~'')
    )%>%
  mutate(Treatment = case_when(
    (EVT_type %in% 'health')~str_c(draxxin, baytril, excenel, ampicillin, excede, nuflor, zuprevo, zactran, oxytet, sulf, pen, ban, elect, thera, dex,  sep = ''), 
    TRUE~'NONE'))%>%
  mutate(Treatment = case_when(
    (Treatment %in% '')~'other', 
    TRUE~Treatment
  ))%>%

# test<-remarks%>%
#   filter(Treatment %in% 'Other')%>%
#   group_by(Event, Remark)%>%
#   summarize(ct_row = sum(n()))%>%
#   ungroup()
#%>%
    group_by(Event, Treatment, floor_date_month, total_cowid, total_row)%>%

  #group_by(Event, Remark_simple, floor_date_month, total_cowid, total_row)%>%
  summarize(ct_cowid = n_distinct(cowid), 
            ct_row = sum(n()))%>%
  ungroup()%>%
  mutate(pct = round((ct_cowid/total_cowid)*100, digits = 0))


outcomes<-events2%>%
  #filter(Event %in% 'PNEU')%>%
  filter(event_ct==1)%>%
  group_by(Event, floor_date_month)%>%
  mutate(total_cowid = n_distinct(cowid), 
            total_row = sum(n()))%>%
  ungroup()%>%
  group_by(Event, outcome_basic, floor_date_month, total_cowid, total_row)%>%
  summarize(ct_cowid = n_distinct(cowid), 
            ct_row = sum(n()), 
            list_remarks = paste0(sort(unique(str_sub(Remark, 1,4))), collapse = ','))%>%
  ungroup()%>%
  mutate(pct = round((ct_cowid/total_cowid)*100, digits = 0))
  
ggplot(outcomes%>%
         filter(Event %in% 'PNEU')%>%
         filter(outcome_basic == 'Success'))+
  geom_bar(data = remarks%>%
         filter(Event %in% 'PNEU'), 
           aes(x = floor_date_month, y = pct, fill = Treatment ), stat = 'identity')+
  geom_vline(xintercept = floor_date(events2$date, unit = 'year'), linetype = 'dashed', color = 'grey')+
  geom_hline(yintercept = 70, color = 'grey')+
  geom_text(aes(x = min(events2$floor_date_month, na.rm = T), y = 70, label = 'Target'),vjust = 0, hjust = 0, color = 'grey')+

  geom_point(aes(x = floor_date_month, y = pct))+
  geom_smooth(data = outcomes%>%
         filter(Event %in% c('PNEU'))%>%
         filter(outcome_basic == 'Success')%>%
           filter(floor_date_month<lubridate::ymd('2024-7-1')),
                  aes(x = floor_date_month, y = pct), span = .3, color = 'black')+
  
  #facet_wrap(Event~., scales = 'free')+
  theme_bw()+
  ylim(c(0,100))+
  xlab('Date')+
  labs(title = 'Monthly 1st Treatment Success Rate')
  
  remarks2<-events2%>%
    filter(Event %in% 'PNEU')%>%
    filter(event_ct ==1)%>%
    group_by((Remark))%>%
    summarize(ct = sum(n()))%>%
    ungroup()
  
  table(remarks2$Remark_simple)
  
  sort(unique(remarks2$Remark))

```

# Calf list

```{r}
calf_list<-events2%>%
  mutate(
    assist = case_when(
    (Event=='ASSIST')~date, 
    TRUE~lubridate::mdy('')
    ), 
    died = case_when(
    (Event=='DIED')~date, 
    TRUE~lubridate::mdy('')
    )
    )%>%
  group_by(cowid, ID, EID, bdat)%>%
  summarise(date_in = min(date), 
            date_out = max(date), 
            list_events = paste0(sort(unique(Event)), collapse = ','), 
            assist_date = lubridate::ymd(paste0(unique(assist), collapse = ',' )), 
            died_date = lubridate::ymd(paste0(unique(died), collapse = ','))
            )%>%
  ungroup()%>%
  mutate(dof = as.numeric(date_out-date_in),
         age_assist = assist_date-bdat, 
         age_died = died_date-bdat)%>%
  mutate(age_in = bdat-date_in, 
         age_out = date_out-bdat)%>%
  mutate(cohort_year = year(date_in), 
         cohort_month = floor_date(date_in, unit = 'month'))


```

#health events
```{r}
health_events<-events2%>%
  filter((EVT_type %in% 'health')|(Event %in% 'DIED'))%>%
  group_by(cowid, ID, EID, bdat, Event, max_event_ct)%>%
  summarize(date_start = min(date), 
            date_end = max(date) 
            )%>%
  ungroup()%>%
  mutate(event_durration = date_end-date_start, 
         age_at_event_start = date_start-bdat, 
         age_at_event_end = date_end-bdat)

health_events_start<-health_events%>%
  mutate(EventStart = paste0(Event, '_start'), 
         EventEnd = paste0(Event, '_end'))%>%
    mutate(list_health = paste0(sort(unique(Event)), collapse = ','))%>%
  select(cowid, ID, EID, bdat, list_health, EventStart, date_start)%>%
  pivot_wider(names_from = EventStart, 
              values_from = date_start)
  

table(health_events_start$EventCombos
      )
  
```


```{r}
ggplot(calf_list)+
  geom_density(aes(x = dof))

ggplot(calf_list)+
  geom_density(aes(x = age_in))

ggplot(calf_list)+
  geom_density(aes(x = age_out))+
  coord_cartesian(xlim = c(0, NA))

ggplot(calf_list)+
  geom_density(aes(x = date_in))

ggplot(calf_list)+
  geom_density(aes(x = date_out))

ggplot(calf_list)+
  geom_bar(aes(x = date_in))
```


